---
title: "final_code"
author: "Connor Riordan"
date: "4/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(dplyr)
library(janitor)
library(ggplot2)
library(gt)
library(skimr)
library(tidyr)
library(broom)
library(lubridate)
library(patchwork)
library(scales)
library(markdown)
library(usmap)
library(readxl)
```

## Presidential Election Results by County 2000-2016

```{r}
pres_results <- read.csv("raw_data/countypres_2000-2016.csv") %>%
  clean_names() %>%
  select(! version)

pres_results <- pres_results %>%
  mutate(abb_county = paste0(county, ", ", state_po))

dat <- pres_results %>%
  filter(abb_county == "Autauga, AL" | abb_county == "Baldwin, AL")

dat %>%
  filter(year == 2000) %>%
  ggplot(aes(candidate, candidatevotes, fill = abb_county)) + 
  geom_bar(stat = "identity", position = position_dodge())

us_counties <- map_data("county")

us_counties

pres_results$region <- tolower(pres_results$state)

newdata <- vector()
for(i in unique(pres_results$state)){
  sub <- pres_results[pres_results$state == i, ]
  for(j in unique(sub$county)){
    sub2 <- sub[sub$county == j, ]
    sub2$candidate <- paste(gsub(' ', '_', sub2$candidate), sub2$year, sep = '_')
    out <- sub2$candidatevotes
    out <- c(i, j, sub2$candidatevotes)
    names(out) <- c('state', 'county', sub2$candidate)
    newdata <- rbind(newdata, out)
  }
}

newdata <- as.data.frame(newdata)
newdata[, 1:2] <- apply(newdata[, 1:2], 2, tolower)

colnames(us_counties)[5:6] <- c('state', 'county')

new_counties <- vector()

for(i in unique(us_counties$state)){
  sub <- us_counties[us_counties$state == i, ]
  for(j in unique(sub$county)){
    sub2 <- sub[sub$county == j, c(1:4)]
    out <- c(i, j, as.list(sub2))
    new_counties <- rbind(new_counties, out)
  }
}

new_counties <- as.data.frame(new_counties)
colnames(new_counties)[1:2] <- c('state', 'county')

newtest <- merge.data.frame(newdata, new_counties, by = c('state', 'county'))
```

```{r}
pres_results_reg <- pres_results %>%
  select(! c(candidate, party, candidatevotes)) %>%
  distinct() %>%
  filter(year == 2016)

pres_results_reg$county <- as.character(pres_results_reg$county)

```

## Income by County

```{r}
income_county <- read_excel("raw_data/per_capita_income_by_county.xlsx") %>%
  clean_names()

income_county <- income_county %>%
  dplyr::rename(
    "county" = table_1_per_capita_personal_income_by_county_2016_2018,
    "personal_income_2016" = x2,
    "personal_income_2017" = x3,
    "personal_income_2018" = x4,
    "personal_income_rank_state_2018" = x5,
    "pct_change_2017" = x6,
    "pct_change_2018" = x7,
    "rank_in_state_pct_change" = x8
  )

income_county <- income_county[6:3223, 1:5]



```


```{r, join}

income_county <- income_county %>%
  mutate(ID = seq.int(nrow(income_county)))

income_county <- income_county %>%
  mutate(state_po = ifelse(income_county$ID <= 68, "AL", 
                             ifelse(between(ID, 70, 99), "AK", 
                                    ifelse(between(ID, 102, 116), "AZ",
                                           ifelse(between(ID, 120, 193), "AR",
                                                  ifelse(between(ID, 196, 253), "CA",
                                                         ifelse(between(ID, 256, 319), "CO",
                                                                ifelse(between(ID, 322, 329), "CT",
                        ifelse(between(ID, 332, 334), "DE",
                               ifelse(between(ID, 336, 336), "DC",
                                      ifelse(between(ID, 339, 405), "FL",
                                             ifelse(between(ID, 408, 556), "GA",
                                                    ifelse(between(ID, 570, 572), "HI",
                                                           ifelse(between(ID, 575, 618), "ID",
                                                                  ifelse(between(ID, 621, 722), "IL",
                                                                         ifelse(between(ID, 725, 816), "IN",
                                                                                ifelse(between(ID, 819, 917), "IA",
                          ifelse(between(ID, 920, 1024), "KS",
                                 ifelse(between(ID, 1027, 1146), "KY",
                                        ifelse(between(ID, 1149, 1212), "LA",
                                               ifelse(between(ID, 1215, 1230), "MD",
                                                      ifelse(between(ID, 1233, 1256), "MA",
                                                             ifelse(between(ID, 1275, 1357), "MI",
                                                                    ifelse(between(ID, 1360, 1446), "MN",
                                                                           ifelse(between(ID, 1449, 1530), "MS",
                                                                                  ifelse(between(ID, 1533, 1647), "MO",
                            ifelse(between(ID, 1650, 1705), "MT",
                                   ifelse(between(ID, 1708, 1800), "NE",
                                          ifelse(between(ID, 1803, 1819), "NV",
                                                 ifelse(between(ID, 1822, 1831), "NH",
                                                        ifelse(between(ID, 1834, 1854), "NJ",
                                                               ifelse(between(ID, 1857, 1889), "NM",
                                                                      ifelse(between(ID, 1892, 1953), "NY",
                                                                             ifelse(between(ID, 1956, 2055), "NC",
                                                                                    ifelse(between(ID, 2058, 2110), "ND",
                              ifelse(between(ID, 2113, 2200), "OH",
                                     ifelse(between(ID, 2203, 2279), "OK",
                                            ifelse(between(ID, 2282, 2317), "OR",
                                                   ifelse(between(ID, 2320, 2386), "PA",
                                                          ifelse(between(ID, 2389, 2393), "RI",
                                                                 ifelse(between(ID, 2396, 2441), "SC",
                                                                        ifelse(between(ID, 2444, 2509), "SD",
                                                                               ifelse(between(ID, 2512, 2606), "TN",
                                                                                      ifelse(between(ID, 2609, 2862), "TX",
                                ifelse(between(ID, 2865, 2893), "UT",
                                       ifelse(between(ID, 2896, 2909), "VT",
                                              ifelse(between(ID, 2912, 2994), "VA",
                                                     ifelse(between(ID, 3021, 3059), "WA",
                                                            ifelse(between(ID, 3062, 3116), "WV", "NA")))))))))))))))))))))))))))))))))))))))))))))))))

income_county$state_po[3119:3190] <- "WI"

income_county$state_po[3193:3215] <- "WY"

income_county_ns <- income_county %>%
  drop_na() %>%
  mutate(abb_county = paste0(county, ", ", state_po))

#income_county_ns <- income_county_ns[-c(2906, 2908, 2910, 2909, 2912, 2913, 2914), ]

income_county_ns <- income_county_ns %>%
  mutate(ID_2 = seq.int(nrow(income_county_ns)))

ic_2016 <- income_county_ns %>%
  select(county, personal_income_2016) %>%
  rename(
    "y" = personal_income_2016
  )

ic_2017 <- income_county_ns %>%
  select(county, personal_income_2017) %>%
  rename(
    "y" = personal_income_2017
  )

ic_2018 <- income_county_ns %>%
  select(county, personal_income_2018) %>%
  rename(
    "y" = personal_income_2018
  )

ic_rank_2018 <- income_county_ns %>%
  select(county, personal_income_rank_state_2018) %>%
  rename(
    "y" = personal_income_rank_state_2018
  )

ic_2016 %>%
  filter(county == "Autauga") %>%
  ggplot(aes(county, y)) + geom_line()
                                                                               

pres_income <- inner_join(pres_results_reg, income_county_ns)

pres_income %>%
  filter(state == "Alabama") %>%
  ggplot(aes(personal_income_2016, totalvotes)) + geom_point()

# ggplot(regression, aes(personal_income_2016, totalvotes, color = county)) + geom_col()
  

```

## County Population Estimate

```{r}
county_pop <- read_csv("raw_data/estimated_county_population.csv") %>%
  clean_names()

county_pop <- county_pop %>%
  select(c(ctyname, popestimate2016))

county_pop <- county_pop[-c(329, 560),] #2867, 2868, 2871, 2878, 2880, 2887, 2889, 2894, 2897, 2901, 2904,
                            #2908, 2910, 2919, 2923, 2926, 2927, 2928, 2929, 2934, 2935),]

county_pop <- county_pop %>%
  mutate(ID_2 = seq.int(nrow(county_pop))) %>%
  dplyr::rename("county_c" = ctyname)
  

pres_income <- pres_income %>%
  mutate(county_c = paste0(county, " County"))

regression <- inner_join(county_pop, income_county_ns)


```

 filter(! county == "Alabama" & ! county == "Alaska" & ! county == "Arizona" &
           ! county == "Arkansas" & ! county == "California" & ! county == "Colorado"  & 
           ! county == "Connecticut" & ! county == "Delaware" & ! county == "Florida" &
           ! county == "Georgia" & ! county == "Hawaii" & ! county == "Idaho" &
           ! county == "Illinois" & ! county == "Indiana" & ! county == "Iowa"  & 
           ! county == "Kansas" & ! county == "Kentucky" & ! county == "Louisiana" & 
           ! county == "Maine" & ! county == "Maryland" & ! county == "Massachusetts" & 
           ! county == "Michigan" & ! county == "Minnesota" & ! county == "Mississippi" & 
           ! county == "Missouri" & ! county == "Montana" & ! county == "Nebraska" & 
           ! county == "Nevada" & ! county == "New Hampshire" & ! county == "New Jersey" & 
           ! county == "New Mexico" & ! county == "New York" & ! county == "North Carolina" & 
           ! county == "North Dakota" & ! county == "Ohio" & ! county == "Oklahoma" & 
           ! county == "Oregon" & ! county == "Pennsylvania" & ! county == "Rhode Island" & 
           ! county == "South Carolina" & ! county == "South Dakota" & ! county =="Tennessee" & 
           ! county == "Texas" & ! county == "Utah" & ! county == "Vermont" & ! county == "Virginia" & 
           ! county == "Washington" & ! county == "West Virginia" & ! county == "Wisconsin" & ! county == "Wyoming"
)

## Health Data (2016)

```{r}
health <- read.csv("raw_data/health_data_2016.csv") %>%
  clean_names() %>%
  select(! 3:6)



```

## State Education Data 2018

```{r}

education <- read.csv("raw_data/state_education_data_2018.csv") %>%
  clean_names() %>%
  select(! 3:7)

# sapply(education, numeric)

education$school_digger_number_of_stars_elementary <- as.numeric(education$school_digger_number_of_stars_elementary)
education$school_digger_number_of_stars_high <- as.numeric(education$school_digger_number_of_stars_high)
education$school_digger_number_of_stars_middle <- as.numeric(education$school_digger_number_of_stars_middle)
education$school_digger_rank_elementary <- as.numeric(education$school_digger_rank_elementary)
education$school_digger_rank_high <- as.numeric(education$school_digger_rank_high)
education$school_digger_rank_middle <- as.numeric(education$school_digger_rank_middle)
education$school_digger_rank_of_elementary <- as.numeric(education$school_digger_rank_of_elementary)
education$school_digger_rank_of_high <- as.numeric(education$school_digger_rank_of_high)
education$school_digger_rank_of_middle <- as.numeric(education$school_digger_rank_of_middle)
education$number_of_all_students_no_value_means_redacted <- as.numeric(education$number_of_all_students_no_value_means_redacted)

education <- education[2:52,]
education %>% filter(qualifying_name == "Connecticut") %>%
ggplot(aes(qualifying_name, school_digger_number_of_stars_elementary)) + geom_col()

```

## Family Voting Registration by Age and Income 2016

```{r}

age_income_2016_n <- read_excel("raw_data/family_voting_registration_by_age_and_income_2016.xlsx", skip = 5, col_types = "numeric") %>%
  clean_names() %>%
  select(! c(x1, x2))

age_income_2016_c <- read_excel("raw_data/family_voting_registration_by_age_and_income_2016.xlsx", skip = 5) %>%
  clean_names() %>%
  select(c(x1, x2))

age_income_2016 <- cbind(age_income_2016_c, age_income_2016_n)

age_income_2016 <- age_income_2016 %>%
  rename(
    "age_group" = x1,
    "total_income" = x2,
    "total_population" = x3,
    "reported_registered" = reported_registered_5,
    "reported_registered_percent" = x6,
    "reported_not_registered_percent" = x8,
    "no_response_to_registration" = no_response_to_registration1,
    "no_response_to_registration_percent" = x10,
    "reported_voted" = reported_voted_11,
    "reported_voted_percent" = x12,
    "reported_not_voted_percent" = x14,
    "no_response_to_voting" = no_response_to_voting2,
    "no_response_to_voting_percent" = x16,
    "registered_total_percent" = reported_registered_17,
    "voted_total_percent" = reported_voted_18
  )
age_income_2016_18_over <- age_income_2016 %>%
  head(13) %>%
  select(! age_group) %>%
  drop_na()

age_income_2016_18_24 <- age_income_2016[14:25,] %>%
  select(! age_group) %>%
  drop_na()

age_income_2016_25_44 <- age_income_2016[26:37,] %>%
  select(! age_group) %>%
  drop_na()

age_income_2016_45_64 <- age_income_2016[38:49,] %>%
  select(! age_group) %>%
  drop_na()

age_income_2016_65_74 <- age_income_2016[50:61,] %>%
  select(! age_group) %>%
  drop_na()

age_income_2016_75_over <- age_income_2016[62:73,] %>%
  select(! age_group) %>%
  drop_na()

age_income_2016_clean <- age_income_2016[2:166, 2:18]
  
```

## Grad Rates by School District

```{r}
grad_ed <- read_excel("raw_data/grad_rates_school_district.xlsx") %>%
  clean_names()

grad_ed <- grad_ed[12:3344,]

grad_ed <- grad_ed %>%
  dplyr::rename(
    "percent" = data,
    "school_district" = location,
    "year" = time_frame
  ) %>%
  select(c(school_district, year, percent))

grad_ed


```














